---
title: "p8105_hw_3_hg2596"
output: github_document
name: Hilina Giday
date: "2022-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
```

# Problem 2 

```{r}
accel_1 = read_csv("./data/accel_data.csv")%>%
  janitor::clean_names() 
  
```

```{r}

accel_2 =  pivot_longer(
    accel_1, 
    activity_1:activity_1440,
    names_to = "mins",  
    names_prefix = "activity_",
    values_to = "count"
    ) %>%
  mutate(Week_identify = ifelse(day%in%c("Saturday", "Sunday"), 2, 1)) %>%
  mutate(day = fct_relevel(day,"Monday","Tuesday","Wednesday", "Thursday", "Friday","Saturday", "Sunday"))
```

Originally had over 1440 columns, the following variables were: **week** (1-5),**day_ID** (just a count), **day** (Mon-Sun), **activity_1**, **activity_2**...activity count following the patient 24 hours. 

Now, after pivoting longer, we are able to put the counts is **counts** making it one variable,and made a new variable **mins** to kep track which count came from which minute of the week.  

There are now `r nrow(accel_2)` observations and `r ncol(accel_2) `  wows


```{r}
accel_3 = 
accel_2 %>%
  mutate(day = as.factor(day))%>%
  group_by(week,day) %>%
  summarize(total_activity = sum(count))
```

maybe ad false = true so the high can come to the beginning 
```{r}
accel_4 = accel_3 %>%
  pivot_wider(
  names_from = "day", 
  values_from = "total_activity")
```

**Question 1; Part 2**
Any trends apparent? 

A big drop on the weekends.


```{r}
ggplot(accel_2, aes(x = mins, y = count, color = day)) + 
  geom_point()
```

```{r}
ggplot(accel_2, aes(x = mins, y = count, color = day)) + 
  geom_line()
```

**Question 1; Part 3**

Describe any patterns or conclusions 
- Most stay under 2500, two big peaks toward the beginning of the day and end. 
- Which one is better representative, the line or the point? 