---
title: "p8105_hw_3_hg2596"
output: github_document
name: Hilina Giday
date: "2022-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
```

# Problem 2 

```{r}
accel_1 = read_csv("./data/accel_data.csv")%>%
  janitor::clean_names() 
  
```

```{r}

accel_2 =  pivot_longer(
    accel_1, 
    activity_1:activity_1440,
    names_to = "mins",  
    names_prefix = "activity_",
    values_to = "count"
    ) %>%
  mutate(Week_identify = ifelse(day%in%c("Saturday", "Sunday"), 2, 1)) %>%
  mutate(day = fct_relevel(day,"Monday","Tuesday","Wednesday", "Thursday", "Friday","Saturday", "Sunday"))
```

Originally had over 1440 columns, the following variables were: **week** (1-5),**day_ID** (just a count), **day** (Mon-Sun), **activity_1**, **activity_2**...activity count following the patient 24 hours. 

Now, after pivoting longer, we are able to put the counts is **counts** making it one variable,and made a new variable **mins** to kep track which count came from which minute of the week.  

There are now `r nrow(accel_2)` observations and `r ncol(accel_2) `  wows


```{r}
accel_3 = 
accel_2 %>%
  mutate(day = as.factor(day))%>%
  group_by(week,day) %>%
  summarize(total_activity = sum(count))
```

maybe ad false = true so the high can come to the beginning 
```{r}
accel_4 = accel_3 %>%
  pivot_wider(
  names_from = "day", 
  values_from = "total_activity")
```

**Question 1; Part 2**
Any trends apparent? 

A big drop on the weekends.


```{r}
ggplot(accel_2, aes(x = mins, y = count, color = day)) + 
  geom_point()
```

```{r}
ggplot(accel_2, aes(x = mins, y = count, color = day)) + 
  geom_line()
```

**Question 1; Part 3**

Describe any patterns or conclusions 
- Most stay under 2500, two big peaks toward the beginning of the day and end. 
- Which one is better representative, the line or the point? 


# Problem 3

```{r}
data("ny_noaa")
```

```{r}
names(ny_noaa)
head(ny_noaa)

```

```{r}
noaa_2 = ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date,c("A", "B", "C")) %>%
  select("B","C","A", everything()) %>%
  rename(month = "B", day = "C", year = "A") %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(tmax = as.integer(tmax)) %>%
  mutate(tmin = as.integer(tmin)) %>%
  select(id, everything()) 

head(noaa_2)

```

```{r}
noaa_3= 
  noaa_2 %>% 
  mutate(tmax_new = tmax/10) %>%
  mutate(tmin_new = tmin/10) %>%
  mutate(prcp_1 = prcp/10) %>% 
  mutate(prcp_new = prcp_1/1000) %>% 
  mutate(snow_new = snow/1000) %>%
  mutate(snwd_new = snwd/1000) %>% 
  mutate_if(is.double,as.integer)
```


tmax: Maximum temperature (tenths of degrees C) -> C
tmin: Minimum temperature (tenths of degrees C) -> C

prcp: Precipitation (tenths of mm) -> m 
snow: Snowfall (mm) -> m 
snwd: Snow depth (mm) -> m 


```{r}
noaa_3 %>% 
  count(snow_new, name = "n_obs", sort = TRUE)
```


** Question 3; 

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

After cleaning,  **ID**, **date** (month,day,year), **prcp**, **snow**, **snwd**, **tmax** and **tmin**. 

**missing** alot...?

For snowfall, what are the most commonly observed values? Why?

Most common is no snow days, "0". The reason behind that can be because location, not too many snow days. 


```{r}
noaa_3 %>%
  mutate(tmax_new = as.numeric(tmax_new)) %>%
  group_by(id, month) %>%
  filter(month == "Jan"| month == "Jul") %>%
  summarize(avg_tmax = mean(tmax_new)) %>%
  ggplot(aes(x = id, y = avg_tmax)) + geom_point(alpha = .8) + facet_grid(. ~ month)

```

Is there any observational, interpretable structure. Any outliers? 
try box plot? 

```{r}

skimr::skim(noaa_3)
  
```

```{r}
graph = 
  noaa_3 %>% 
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = tmax_new, y = tmin_new)) + 
  geom_hex()
```


```{r}
scatter = 
  noaa_3 %>%
  filter(snow_new > 0) %>%
  filter(snow_new < 100) %>%
  ggplot(aes(x = snow_new , y = year)) + geom_line()
```


no question asked. 

helpful with axis: 

```{r}
#theme(axis.text.x = element_blank(), legend.position = "none")
```























scatter_snow = 
noaa_3 %>% 
  mutate(year = as.factor(year)) %>%
  filter(snow_new > 0) %>%
  filter(snow_new <100) %>%
ggplot(aes(x = snow_new, y = year)) + 
  geom_line() +
  theme(legend.position = "none")

(hex_temp/scatter_snow)














