---
title: "p8105_hw_3_hg2596"
output: github_document
name: Hilina Giday
date: "2022-10-15"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      fig.width = 8,
                      fig.height = 6, 
                      out.width = "90%")
```

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)
```

# Problem 2 

```{r}
accel_1 = read_csv("./data/accel_data.csv")%>%
  janitor::clean_names() 
  
```

```{r}

accel_2 =  pivot_longer(
    accel_1, 
    activity_1:activity_1440,
    names_to = "mins",  
    names_prefix = "activity_",
    values_to = "count"
    ) %>%
  mutate(Week_identify = ifelse(day%in%c("Saturday", "Sunday"), 2, 1)) %>%
  mutate(day = fct_relevel(day,"Monday","Tuesday","Wednesday", "Thursday", "Friday","Saturday", "Sunday"))
```

Originally had over 1440 columns, the following variables were: **week** (1-5),**day_ID** (just a count), **day** (Mon-Sun), **activity_1**, **activity_2**...activity count following the patient 24 hours. 

Now, after pivoting longer, we are able to put the counts is **counts** making it one variable,and made a new variable **mins** to kep track which count came from which minute of the week.  

There are now `r nrow(accel_2)` observations and `r ncol(accel_2) `  wows


```{r}
accel_3 = 
accel_2 %>%
  mutate(day = as.factor(day))%>%
  group_by(week,day) %>%
  summarize(total_activity = sum(count))
```

maybe ad false = true so the high can come to the beginning 
```{r}
accel_4 = accel_3 %>%
  pivot_wider(
  names_from = "day", 
  values_from = "total_activity")

accel_4 
```

**Question 1; Part 2**
Any trends apparent? 

A big drop on the weekends.

Just eyeballing the table, it looks like the for week 4 and week 5 there is a **big drop** in counts on Saturday. On average, Mondays on average higher counts. Overall, though its difficult to see very apparent trends just based on the table. 

```{r}
ggplot(accel_2, aes(x = mins, y = count, color = day)) + 
  geom_line()
```

**Question 1; Part 3**

Describe any patterns or conclusions 

There is a lot of data, so it is very difficult to distinguish patterns.However, some of the trends is that that count on **average is below 2500**. Another trend is there seems to be peak minutes when the count goes above the average count, its difficult to know exactly what exact minutes but it seems their are **two high peaks** during the morning and later on toward the latter portion of the 24 hour period. 


# Problem 3

```{r}
data("ny_noaa")
```

```{r}
names(ny_noaa)
head(ny_noaa)

```

```{r}
noaa_2 = ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date,c("A", "B", "C")) %>%
  select("B","C","A", everything()) %>%
  rename(month = "B", day = "C", year = "A") %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(tmax = as.integer(tmax)) %>%
  mutate(tmin = as.integer(tmin)) %>%
  select(id, everything()) 

head(noaa_2)

```

```{r}
noaa_3= 
  noaa_2 %>% 
  mutate(tmax_new = tmax/10) %>%
  mutate(tmin_new = tmin/10) %>%
  mutate(prcp_new = prcp/10) %>% 
  mutate_if(is.double,as.integer)
```


tmax: Maximum temperature (tenths of degrees C) -> C
tmin: Minimum temperature (tenths of degrees C) -> C
prcp: Precipitation (tenths of mm) -> m 

```{r}
noaa_3 %>% 
  count(snow_new, name = "n_obs", sort = TRUE)
```


** Question 3; 

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

After cleaning,  **ID**, **date** (month,day,year), **prcp**, **snow**, **snwd**, **tmax** and **tmin**. 

**missing** alot...?

For snowfall, what are the most commonly observed values? Why?

Most common is no snow days, "0". The reason behind that can be because location, not too many snow days. 

*Is missing data an issue??????????


```{r}
noaa_3 %>%
  mutate(tmax_new = as.numeric(tmax_new)) %>%
  group_by(id, month) %>%
  filter(month == "Jan"| month == "Jul") %>%
  summarize(avg_tmax = mean(tmax_new)) %>%
  ggplot(aes(x = id, y = avg_tmax)) + geom_point(alpha = .8) + facet_grid(. ~ month)

```

Is there any observational, interpretable structure. Any outliers? 
try box plot? 

```{r}

skimr::skim(noaa_3)
  
```

```{r}
graph = 
  noaa_3 %>% 
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = tmax_new, y = tmin_new)) + 
  geom_hex()
```


```{r}

scatter_2 = 
  noaa_2 %>%
  mutate(year = as.factor(year)) %>%
  filter(snow < 100) %>%
  filter(snow >= 0) %>%
  ggplot(aes(x = snow , y = year)) + geom_density_ridges(alpha = 0.5)

```

```{r}
(graph/scatter_2)
```

ADD titles??




















scatter_snow = 
noaa_3 %>% 
  mutate(year = as.factor(year)) %>%
  filter(snow_new > 0) %>%
  filter(snow_new <100) %>%
ggplot(aes(x = snow_new, y = year)) + 
  geom_line() +
  theme(legend.position = "none")

(hex_temp/scatter_snow)














